---
title: "Data Analysis HW 1"
author: "Sina Alemohammad - 93111706"
date: "February 11, 2018"
output: html_document
---
## Problem 1

The known Magic matrices are a special case of the "friendship" matrices which the sum of the all rows, columns and diagonal of the matrix is the same. But in this problem only sum of the rows and columns should be the same, so any permutations of the rows and columns of the magic matrices is a possible answer. In this report, an example is provided for each n = 4,5,6.

```{r friendship, message=FALSE, warning=FALSE}
library(magic)
m4 <- magic(4)
m5 <- magic(5)
m6 <- magic(6)
e <- sample(x = 1:4 ,size = 4 ,replace = FALSE)
example.n4 = m4[e,] 
example.n4
e <- sample(x = 1:5 ,size = 5 ,replace = FALSE)
example.n5 = m5[e,] 
example.n5
e <- sample(x = 1:6 ,size = 6 ,replace = FALSE)
example.n6 = m6[e,] 
example.n6
```

## Problem 2

We have to consider all the permutations of 1 to 13 and check if each permutation satisfies the conditions but calculating all the permutations and checking all them demands too much memory and time. As there are 3 sets of conditions, we can break the problem into different parts and look for valid permutations of each part. So I broke the problem into two parts, first by calculating combination of 6 out of 13, the valid sequences that satisfies the second conditions, 26 sequence in this problem, are found and then all the other possible permutations of the remaining numbers are calculated and tested to find which sequence, only one sequence in this problem, satisfies all the conditions.  
```{r strange sequence ,message=FALSE, warning=FALSE}
library(arrangements)
y = matrix(0,nrow = 1, ncol = 6)[FALSE,]
condition2.possible.combinations = permutations(13,6)
for (i in 1:nrow(condition2.possible.combinations)){
  x <-  condition2.possible.combinations[i,]
  c3 <- x[2]/(x[2]+x[3])
  c4 <- (x[2]+x[3])/(x[3]+x[4]+x[5])
  c5 <- (x[3]+x[4]+ x[5])/(x[1]+x[5]+x[6])
  if (c3 == c4 & c3 == c5){
    y = rbind(y,x)
  }
}

P <- permutations(7)
n <- 1:13
for (i in 1:nrow(y)){
  temp = n[!n %in% y[i,]]
  for (j in 1:nrow(P)){
    x <- c(temp[P[j,1:4]], y[i,],temp[P[j,5:7]])
    c1 <- x[1]/(x[2]+x[3])
    c2 <- (x[2]+x[3])/(x[3]+x[4]+x[5])
    c6 <- (x[11]+x[12])/(x[12]+x[13])
    c7 <- (x[12]+x[13])/(x[13]+x[10])
    if (c1 == c2 & c6 == c7 & c1 < 1 & c6 < 1){
      print(x)
    }
  }
}
```

## Problem 3 

For all parts of this problem, all the possible combinations of liars and trustfuls are calculated and tested to see if it satisfies the conditions and then the number of liars for the possible combinations are announced.

+ Part 1
```{r  ,message=FALSE, warning=FALSE}

n <- 16
l <- rep(list(0:1), n)
x = expand.grid(l)
Number.of.liers = as.matrix(0,nrow = 0,ncol = 0)
Number.of.liers = Number.of.liers[FALSE]
for (i in 1:2^n){
  temp = x[i,]
  w = which(temp == 0)
  flag = 0
  if (length(w) != 0){
  for (j in 1:length(w)){
    a = (w[j]-2):w[j]
    if (sum(temp[a%%n+1]) != 2) {
      flag = flag + 1
      break
    }
  }
  }
    w = which(temp == 1)
    if (flag == 0){
    if (length(w) != 0){
    for (j in 1:length(w)){
      a = (w[j]-2):w[j]
      if (sum(temp[a%%n+1]) == 3) {
        flag = flag + 1
        break
      }
    }
    }
    }
    if (flag == 0){
      Number.of.liers = c(Number.of.liers,sum(temp))
    }
}
unique(Number.of.liers)
```

+ Part 2
```{r  ,message=FALSE, warning=FALSE}
n <- 12
l <- rep(list(0:1), n)
x = expand.grid(l)
Number.of.liers = as.matrix(0,nrow = 0,ncol = 0)
Number.of.liers = Number.of.liers[FALSE]
for (i in 1:2^n){
  temp = x[i,]
  w = which(temp == 0)
  flag = 0
  if (length(w) != 0){
    for (j in 1:length(w)){
      a = (w[j]-2):w[j]
      if (sum(temp[a%%n+1]) != 1) {
        flag = flag + 1
        break
      }
    }
  }
  w = which(temp == 1)
  if (flag == 0){
    if (length(w) != 0){
      for (j in 1:length(w)){
        a = (w[j]-2):w[j]
        if (sum(temp[a%%n+1]) == 2) {
          flag = flag + 1
          break
        }
      }
    }
  }
  if (flag == 0){
    Number.of.liers = c(Number.of.liers,sum(temp))
  }
}
unique(Number.of.liers)
```

+ Part 3
```{r  ,message=FALSE, warning=FALSE}
n <- 8
l <- rep(list(0:1), n)
x = expand.grid(l)
Number.of.liers = as.matrix(0,nrow = 0,ncol = 0)
Number.of.liers = Number.of.liers[FALSE]
for (i in 1:2^n){
  for(k in 1:2^n){
    if (sum(x[k,]) == 4){
      temp = x[i,]
      temp2 = x[k,]
      w = which(temp == 0)
      flag = 0
      if (length(w) != 0){
        for (j in 1:length(w)){
          a = (w[j]-2):w[j]
          if (sum(temp[a%%n+1]) != 1 & temp2[w[j]] == 1 ) {
            flag = flag + 1
            break
          }
          if (sum(temp[a%%n+1]) != 2 & temp2[w[j]] == 0 ) {
            flag = flag + 1
            break
          }
        }
      }
      w = which(temp == 1)
      if (flag == 0){
        if (length(w) != 0){
          for (j in 1:length(w)){
            a = (w[j]-2):w[j]
            if (sum(temp[a%%n+1]) == 2 & temp2[w[j]] == 1 ) {
              flag = flag + 1
              break
            }
            if (sum(temp[a%%n+1]) == 3 & temp2[w[j]] == 0 ) {
              flag = flag + 1
              break
            }
          }
        }
      }
      if (flag == 0){
        Number.of.liers = c(Number.of.liers,sum(temp))
      }
    }
  }
}
unique(Number.of.liers)

```